---
title: "Project Data Analysis"
author: "Anthony Bruno"
date: "2023-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
```

Load raw Data
```{r}
# Specify the file path to your CSV file
file_path <- "rawdata.csv"

# Read the CSV file and create a data frame
streamData <- read.csv(file_path)
```

Visualize the data:
 
Look at Total N First:
```{r}
#Look at Total N vs. Treatment Type
ggplot(data = streamData, aes(x= Type, y = Total.N)) +
  stat_boxplot() + 
  ggtitle("Total N for each treatment") 

#Plot of Total N vs. Distance from Stream Bank
ggplot( data = streamData, aes( x = Distance, y = Total.N, color = Type, shape = Type)) + 
  geom_point() +
  ggtitle("Total N vs. Distance from Stream Bank") +
  theme_minimal()

#Box plots of Total N for each type at each distance
ggplot(streamData, aes( y = Total.N, fill = Type )) +
  geom_boxplot() +
  facet_wrap(~ Distance, nrow = 1) +
  labs( x = "Distance from stream bank") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Total N concentrations at Each Distance") + 
  scale_fill_manual(values = c("CSS" = "red", "No CSS" = "blue"))

#Violin plots of Total N for each type at each distance
ggplot(streamData, aes( x = Distance, y = Total.N, fill = Type)) +
  geom_violin() +
  facet_wrap(~ Distance, nrow = 1) +
  labs( x = "Distance from Stream Bank", y = "Total N Concentration [mg/L]") +
 theme(axis.text.x = element_blank()) +
  ggtitle("Total N concentrations at Each Sampling Distance") +
  scale_fill_manual(values = c("CSS" = "red", "No CSS" = "blue"))
```

Visualize Total C data:

```{r}
#Total C concentrations for each Treatment Type
ggplot(data = streamData, aes(x= Type, y = Total.C)) +
  stat_boxplot() + 
  ggtitle("Total C for each treatment")

#Plot of Inorganic N vs distance colored by treatment type
ggplot( data = streamData, aes( x = Distance, y = Total.C, color = Type)) + 
  geom_point() +
  ggtitle("Total C vs. Distance from Stream Bank")

#Box plot of Inorganic N vs Treatment type at each distance
ggplot(streamData, aes( y = Total.C, fill = Type)) +
  geom_boxplot() +
  facet_wrap(~ Distance, nrow = 1) +
  labs( x = "Distance from stream bank") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Total C concentrations at Each Sampling Distance") +
  scale_fill_manual(values = c("CSS" = "red", "No CSS" = "blue"))

#Violin plots of Total N for each type at each distance
ggplot(streamData, aes( x = Distance, y = Total.C, fill = Type)) +
  geom_violin() +
  facet_wrap(~ Distance, nrow = 1) +
  labs( x = "Distance from Stream Bank", y = "Inorganic N Concentration [mg/L]") +
 theme(axis.text.x = element_blank()) +
  ggtitle("Total C concentrations at Each Sampling Distance") +
  scale_fill_manual(values = c("CSS" = "red", "No CSS" = "blue"))
```
Visualize Elevation Data

```{r}
#Relative Elevation for each Treatment Type
ggplot(data = streamData, aes(x= Type, y = Relative.Elev.)) +
  stat_boxplot() + 
  ggtitle("Total C for each treatment")

#Plot Elevation points
ggplot(streamData, aes( x= Distance, y = Relative.Elev., color = Type)) +
geom_point() +
ggtitle("Relative Elevation vs. Distance from Stream Bank")

#Box plot of Elevation vs. Treatment type at each distance
ggplot(streamData, aes( y = Relative.Elev., fill = Type)) +
  geom_boxplot() +
  facet_wrap(~ Distance, nrow = 1) +
  labs( x = "Distance from stream bank") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Sample Elevations vs. Each Distance") + 
  scale_fill_manual(values = c("CSS" = "red", "No CSS" = "blue"))

ggplot(streamData, aes( x = Distance, y = Relative.Elev., fill = Type)) +
  geom_violin() +
  facet_wrap(~ Distance, nrow = 1) +
  labs( x = "Distance from Stream Bank") +
  theme(axis.text.x = element_blank()) +
  ggtitle("Sample Elevations vs. Each Distance") + 
  scale_fill_manual(values = c("CSS" = "red", "No CSS" = "blue"))
```

Bootstrapping Statistics:

Set the number of resampling replications for each test
```{r}
num_replications <- 100  # Change this to the desired number of replications
```

Compare Total N Concentration for CSS vs non-CSS:

Create function to calculate and store resampled differences
```{r}
resampled_tn_differences <- function(thisStreamData) {
  # Resample the response variable with replace = TRUE
  thisStreamData$resampledResponse <- sample(thisStreamData$Total.N, replace = TRUE)
  
  # Create a new data frame to store the resampled results
  resampled_data <- data.frame(Type = thisStreamData$Type, resampledResponse = thisStreamData$resampledResponse)
  
  # Calculate the Average Total N for each treatment type in the resampled data
  summed_data <- resampled_data %>%
    group_by(Type) %>%
    summarize(tot_n = mean(resampledResponse))
  
  # Calculate the difference in total N for each treatment type
  diff_data <- summed_data %>%
    mutate(tot_n_diff = diff(tot_n))
  
  return(diff_data$tot_n_diff[1])
}

#replicate and store differences
resampled_tn_diff <- replicate(num_replications, resampled_tn_differences(streamData))

#Calculate the observed difference
observed_tn_diff <- streamData %>%
  group_by(Type) %>%
  summarize(avg_tn = mean(Total.N)) %>%
  summarize(observed_diff = diff(avg_tn))
#Calculate the p-value of observed TN diff > resampled differences for total nitrogen concentration among all samples
P_tn_diff <- mean(resampled_tn_diff > observed_tn_diff$observed_diff)

message("The average Total Nitrogen concentrations from CSS samples is", observed_tn_diff$observed_diff, " [mg/L] higher/lower than the Total Nitrogen Concentrations from NCSS samples. Bootstrapping analysis P-value of " , P_tn_diff)
```

Create distribution of Tot N differences (should be approximately normal)
```{r}
hist(resampled_tn_diff)
```

Analyze the difference between TN concentrations at a single sampling distance and check the significance of the observed difference
```{r}
#Specify distance as dist
dist_n <- 2.5

resampled_tn_dist_difference <- function(thisStreamData) {
 thisStreamData %>% 
    filter(Distance == dist_n)
  
    # Resample the response variable with replace = TRUE
  thisStreamData$resampled_tn <- sample(thisStreamData$Total.N, replace = TRUE)
  
  # Create a new data frame to store the resampled results
  resampled_tn_inc <- data.frame(Type = thisStreamData$Type, resampledResponse = thisStreamData$resampled_tn)
  
  # Calculate the average elevation above the WSE for each treatment type in the resampled data
  summed_data <- resampled_tn_inc %>%
    group_by(Type) %>%
    summarize(avg_tn = mean(resampledResponse))
  
  # Calculate the difference in average elevation for each treatment type
  diff_data <- summed_data %>%
    mutate(avg_tn_diff = (diff(avg_tn)))  
  
  return(diff_data$avg_tn_diff[1])
}

#Replicate and store differences
resampled_tn_dist_diff <- replicate(num_replications, resampled_tn_dist_difference(streamData))

#Find Probability of differences greater than or equal too observed_tn_inc_diff
observed_tn_dist_diff <- streamData %>%
  filter(Distance == dist_n) %>%
  group_by(Type) %>%
  summarize(avg_tn = mean(Total.N)) %>%
  summarize(observed_diff = diff(avg_tn))

P_tn_dist_diff <- mean(resampled_tn_dist_diff > observed_tn_dist_diff$observed_diff)

message("The difference of the sum of Total Nitrogen concentrations at a sample distance of ", dist_n," feet is ", observed_tn_dist_diff$observed_diff," [mg/L]. Bootstrapping analysis yielded a P-value of ", P_tn_dist_diff)
```

```{r}
hist(resampled_tn_dist_diff)
```

Look at the probability that there is a linear relationship between all TOTAL NITROGEN samples with respect to DISTANCE from stream bank:

Function for resampling and fitting a linear relationship to the data
```{r}
resampleAndFit_totn_dist=function(thisStreamData){
  #within the function, streamData does not exist, but thisStreamData does
  
  #resample the response variable with replace = true
  thisStreamData$resampledResponse=sample(thisStreamData$Total.N,replace=T)
  
  #fit a lm to the resampled dataset, save it as resampleFit
  resampleFit_totn_dist=lm(thisStreamData$resampledResponse~thisStreamData$Distance)
  
  #points or fitted lines representing the resampled data can be added to the plot by uncommenting out these lines:
  #points(resampledResponse~thisStreamData$Distance,pch=2)
  #abline(resampleFit)
  
  #return the slope coefficient
 return(resampleFit_totn_dist$coefficients[[2]])
}

#Resample the random slopes XXX times
resampledSlopes_totn_dist=replicate(num_replications,resampleAndFit_totn_dist(streamData))

#Calculate observed slope
observedFit_totn_dist=lm(streamData$Total.N~streamData$Distance)

#P-value that that there is a linear relationship between Total N and distance from the stream channel with a slope of

P_tn_slope_dist <- mean(resampledSlopes_totn_dist> observedFit_totn_dist$coefficients[[2]])

message("A linear fit of Total Nitrogen vs. Distance from the wetted bank yielded a slope of ", observedFit_totn_dist$coefficients[[2]]," mg/L/ft. Bootstrapping analysis yielded a P-value of ", P_tn_slope_dist," that the slope is greater than or equal to the observed slope.")

#plot linear fit
ggplot(streamData, aes(x= Distance, y = Total.N, color = Type)) +
  geom_point() + 
  geom_smooth(method = 'lm')
```

Look at the distribution of the resampled slopes:
```{r}
hist(resampledSlopes_totn_dist)
```

Analyze Total C Data:

Create function to calculate and store Total C resampled differences
```{r}
resampled_tc_diff <- function(thisStreamData) {
  # Resample the response variable with replace = TRUE
  thisStreamData$resampledResponse <- sample(thisStreamData$Total.C, replace = TRUE)
  
  # Create a new data frame to store the resampled results
  resampled_data <- data.frame(Type = thisStreamData$Type, resampledResponse = thisStreamData$resampledResponse)
  
  # Calculate the Average Total C for each treatment type in the resampled data
  summed_data <- resampled_data %>%
    group_by(Type) %>%
    summarize(tot_c = mean(resampledResponse))
  
  # Calculate the difference in total N for each treatment type
  diff_data <- summed_data %>%
    mutate(tot_c_diff = diff(tot_c))
  
  return(diff_data$tot_c_diff[1])
}

#replicate and store differences
resampled_tc_diff <- replicate(num_replications, resampled_tc_diff(streamData))

#Calculate the observed difference
observed_tc_diff <- streamData %>%
  group_by(Type) %>%
  summarize(avg_tc = mean(Total.C)) %>%
  summarize(observed_diff = diff(avg_tc))

#Calculate the p-value of observed TC diff > resampled differences for Total Carbon concentration among all samples
P_tc_diff <- mean(resampled_tc_diff > observed_tc_diff$observed_diff)

message("The sum of the Total carbon concentrations from CSS samples is ", observed_tc_diff$observed_diff, " is [mg/L] higher/lower than the Total Carbon concentrations from NCSS samples. Bootstrapping analysis yielded a P-value of ", P_tc_diff)

```

Analyze the difference between TC concentrations at a single sampling distance and check the significance of the observed difference
```{r}
#Specify distance as dist_c
dist_c <- 2.5

resampled_tc_dist_difference <- function(thisStreamData) {
 thisStreamData %>% 
    filter(Distance == dist_c)
  
    # Resample the response variable with replace = TRUE
  thisStreamData$resampled_tc <- sample(thisStreamData$Total.C, replace = TRUE)
  
  # Create a new data frame to store the resampled results
  resampled_tc_inc <- data.frame(Type = thisStreamData$Type, resampledResponse = thisStreamData$resampled_tc)
  
  # Calculate the average elevation above the WSE for each treatment type in the resampled data
  summed_data <- resampled_tc_inc %>%
    group_by(Type) %>%
    summarize(avg_tc = mean(resampledResponse))
  
  # Calculate the difference in average elevation for each treatment type
  diff_data <- summed_data %>%
    mutate(avg_tc_diff = (diff(avg_tc)))  
  
  return(diff_data$avg_tc_diff[1])
}

#Replicate and store differences
resampled_tc_dist_diff <- replicate(num_replications, resampled_tc_dist_difference(streamData))

#Find the difference between TC at CSS vs. NCSS sites at distance = dist
observed_tc_dist_diff <- streamData %>%
  filter(Distance == dist_c) %>%
  group_by(Type) %>%
  summarize(avg_tc = mean(Total.C)) %>%
  summarize(observed_diff = diff(avg_tc))

#Find Probability of differences greater than or equal too observed_tn_inc_diff
P_tc_dist_diff <- mean(resampled_tc_dist_diff > observed_tc_dist_diff$observed_diff)

message("The difference of the sum of Total Carbon concentrations at a sample distance of ", dist_c," feet is ", observed_tc_dist_diff$observed_diff," [mg/L]. Bootstrapping analysis yielded a P-value of ", P_tc_dist_diff)
```

Look at the probability that there is a linear relationship between all Total Carbon samples with respect to DISTANCE from stream bank:

Function for resampling and fitting a linear relationship to the data
```{r}
resampleAndFit_inorgn_dist=function(thisStreamData){
  #within the function, streamData does not exist, but thisStreamData does
  
  #resample the response variable with replace = true
  thisStreamData$resampledResponse=sample(thisStreamData$Total.C,replace=T)
  
  #fit a lm to the resampled dataset, save it as resampleFit
  resampleFit_inorgn_dist=lm(thisStreamData$resampledResponse~thisStreamData$Distance)
  
  #points or fitted lines representing the resampled data can be added to the plot by uncommenting out these lines:
  #points(resampledResponse~thisStreamData$Distance,pch=2)
  #abline(resampleFit)
  
  #return the slope coefficient
 return(resampleFit_inorgn_dist$coefficients[[2]])
}

#Replicate and store
resampledSlopes_inorgn_dist=replicate(num_replications,resampleAndFit_inorgn_dist(streamData))
 
#Calculate observed slope of TC vs. distance
observedFit_totc_dist=lm(streamData$Total.C~streamData$Distance)

#P-value that that there is a linear relationship between Total N and distance from the stream channel with a slope of
P_tc_slope_dist <- mean(resampledSlopes_inorgn_dist > observedFit_totc_dist$coefficients[[2]])

message("A linear fit of Total Carbon vs. Distance from the wetted bank yielded a slope of ", observedFit_totc_dist$coefficients[[2]]," mg/L/ft. Bootstrapping analysis yielded a P-value of ", P_tc_slope_dist, " that the slope is greater than or equal to the observed slope")

#plot linear fits
ggplot(streamData, aes(x= Distance, y = Total.C, color = Type)) +
  geom_point() + 
  geom_smooth(method = 'lm')
```

Look at the distribution of the resampled slopes:
```{r}
hist(resampledSlopes_inorgn_dist)
```
Analyze Elevation vs. Concentration Data:

Look at the probability that there is a linear relationship between all Total Nitrogen samples with respect to Relative Elevation above the Thalweg:

function for resampling and fitting the data
```{r}
resampleAndFit_totn_Elev=function(thisStreamData){
  #within the function, streamData does not exist, but thisStreamData does
  
  #resample the response variable with replace = true
  thisStreamData$resampledResponse=sample(thisStreamData$Total.N,replace=T)
  
  #fit a lm to the resampled dataset, save it as resampleFit
  resampleFit_totn_Elev=lm(thisStreamData$resampledResponse~thisStreamData$Relative.Elev.)
  
  #points or fitted lines representing the resampled data can be added to the plot by uncommenting out these lines:
  #points(resampledResponse~thisStreamData$Distance,pch=2)
  #abline(resampleFit)
  
  #return the slope coefficient
 return(resampleFit_totn_Elev$coefficients[[2]])
}

#resample and store
resampledSlopes_totn_Elev=replicate(num_replications,resampleAndFit_totn_Elev(streamData))

#Calculate Observed Slope of TN vs. Elev.
observedFit_totn_elev = lm(streamData$Total.N~streamData$Relative.Elev.)

#P-value that that there is a linear relationship between Total N and Relative Elevation above the Thalweg
P_tn_slope_elev <- mean(resampledSlopes_totn_Elev > observedFit_totn_elev$coefficients[[2]])

message("A linear fit of Total Nitrogen vs. Elevation from the wetted bank yielded a slope of ", observedFit_totn_elev$coefficients[[2]]," mg/L/ft. Bootstrapping analysis yielded a P-value of ", P_tn_slope_elev, " that the slope is greater than or equal to the observed slope")

#plot linear fit
ggplot(streamData, aes(x= Distance, y = Relative.Elev., color = Type)) +
  geom_point() + 
  geom_smooth(method = 'lm')
```

```{r}
hist(resampledSlopes_totn_Elev)
```

Look at the probability that there is a linear relationship between all Total Carbon samples with respect to Relative Elevation above the Thalweg:

function for resampling and fitting the data
```{r}
resampleAndFit_tc_Elev=function(thisStreamData){
  #within the function, streamData does not exist, but thisStreamData does
  
  #resample the response variable with replace = true
  thisStreamData$resampledResponse=sample(thisStreamData$Total.C,replace=T)
  
  #fit a lm to the resampled dataset, save it as resampleFit
  resampleFit_tc_Elev=lm(thisStreamData$resampledResponse~thisStreamData$Relative.Elev.)
  
  #points or fitted lines representing the resampled data can be added to the plot by uncommenting out these lines:
  #points(resampledResponse~thisStreamData$Distance,pch=2)
  #abline(resampleFit)
  
  #return the slope coefficient
 return(resampleFit_tc_Elev$coefficients[[2]])
}

#replicate and store
resampledSlopes_tc_Elev = replicate(num_replications,resampleAndFit_tc_Elev(streamData))

#Calculate Observed Slope of TN vs. Elev.
observedFit_tc_elev = lm(streamData$Total.C~streamData$Relative.Elev.)

#P-value that that there is a linear relationship between Total C and Relative Elevation above the Thalweg
P_tc_slope_elev <- mean(resampledSlopes_tc_Elev > observedFit_tc_elev$coefficients[[2]])

message("A linear fit of Total Carbon vs. Elevation from the wetted bank yielded a slope of ", observedFit_tc_elev$coefficients[[2]]," mg/L/ft. Bootstrapping analysis yielded a P-value of ", P_tc_slope_elev, " that the slope is greater than or equal to the observed slope")

#plot linear fit
ggplot(streamData, aes(x= Distance, y = Total.C, color = Type)) +
  geom_point() + 
  geom_smooth(method = 'lm')
```

```{r}
hist(resampledSlopes_tc_Elev)
```
Analyze Elevation vs. Treatment:

Look at the probability that there is a difference between the average elevation above the Thalweg for the two treatments:
```{r}
resampled_avgelev_diff <- function(thisStreamData) {
 
    # Resample the response variable with replace = TRUE
  thisStreamData$resampled_elev <- sample(thisStreamData$Relative.Elev., replace = TRUE)
  
  # Create a new data frame to store the resampled results
  resampled_elev_data <- data.frame(Type = thisStreamData$Type, resampledResponse = thisStreamData$resampled_elev)
  
  # Calculate the average elevation above the WSE for each treatment type in the resampled data
  summed_data <- resampled_elev_data %>%
    group_by(Type) %>%
    summarize(avg_elev = mean(resampledResponse))
  
  # Calculate the difference in average elevation for each treatment type
  diff_data <- summed_data %>%
    mutate(avg_elev_diff = (diff(avg_elev)))  # added a default value for the first difference
  
  return(diff_data$avg_elev_diff[1])
}

#replicate and store
resampled_elev_differences <- replicate(num_replications, resampled_avgelev_diff(streamData))

#Calculate the observed average elevation difference for the two treamtments
observed_elev_diff <- streamData %>%
  group_by(Type) %>%
  summarize(avg_elev = mean(Relative.Elev.)) %>%
  summarize(observed_diff = diff(avg_elev))

#Find Probability of differences greater than or equal too observed_elev_diff
P_avg_elev_diff <- mean(resampled_elev_differences > observed_elev_diff$observed_diff)

message("The average relative elevation of surveyed locations without channel spanning structures is ", observed_elev_diff, " feet higher than the average relative elevation of surveyed locations with channel spanning structures. Bootstrapping analysis yielded a low P-value of " , P_avg_elev_diff,  " showing that this observed difference is statistically significant.")

#Create a df showing average relative elevations for the two treatments
Avg_elev <- streamData %>%
  group_by(Type) %>%
  summarize(avg_elev = mean(Relative.Elev.))
```

Create distribution of Average Elev. Differences (should be approximately normal)
```{r}
hist(resampled_elev_differences)
```

Find Probability of differences greater than or equal too observed_elev_diff for specific distances
```{r}
#Specify distance as inc
inc <- 2.5

resampled_dist_elev_diff <- function(thisStreamData) {
 thisStreamData %>% 
    filter(Distance == inc)
  
    # Resample the response variable with replace = TRUE
  thisStreamData$resampled_elev <- sample(thisStreamData$Relative.Elev., replace = TRUE)
  
  # Create a new data frame to store the resampled results
  resampled_single_elev_data <- data.frame(Type = thisStreamData$Type, resampledResponse = thisStreamData$resampled_elev)
  
  # Calculate the average elevation above the WSE for each treatment type in the resampled data
  summed_data <- resampled_single_elev_data %>%
    group_by(Type) %>%
    summarize(avg_elev = mean(resampledResponse))
  
  # Calculate the difference in average elevation for each treatment type
  diff_data <- summed_data %>%
    mutate(avg_elev_diff = (diff(avg_elev)))  
  
  return(diff_data$avg_elev_diff[1])
}

resampled_single_elev_diff <- replicate(num_replications, resampled_dist_elev_diff(streamData))

#Find Probability of differences greater than or equal too observed_single_elev_diff
observed_single_elev_diff <- streamData %>%
  filter(Distance == inc) %>%
  group_by(Type) %>%
  summarize(avg_elev = mean(Relative.Elev.)) %>%
  summarize(observed_diff = diff(avg_elev))

P_avg_single_elev_diff <- mean(resampled_single_elev_diff > observed_single_elev_diff$observed_diff)

message("The average relative elevation " , inc, " feet from the wetted bank is ", observed_single_elev_diff$observed_diff, " feet higher/lower at NCSS sites vs CSS sites. Bootstrapping analysis yielded a P-value of ", P_avg_elev_diff)

```

```{r}
hist(resampled_single_elev_diff)
```

