---
title: "Project Data Analysis"
author: "Anthony Bruno"
date: "2023-10-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
```
#R code is just text - I could share a .r file, but d2l would complain about it because it is unfamiliar with the file type.  Instead, here is the text.  Create a script in R, and past this in.

Load raw Data
```{r}
# Specify the file path to your CSV file
file_path <- "rawdata.csv"

# Read the CSV file and create a data frame
streamData <- read.csv(file_path)
```

Look at Total N vs. Type
```{r}
boxplot(streamData$Total.N~streamData$Type)
```

#plot of Total N vs. distance
```{r}
plot(streamData$Total.N~streamData$Distance)
```

#fit a linear model using lm(), and the same formula used to make the plot.  Save the lm as 'fit'
```{r}
fit=lm(streamData$Total.N[streamData$Type=="CSS"]~streamData$Distance[streamData$Type=="CSS"])
```

#add the fitted line to the plot using the lm object 'fit'
```{r}
plot(streamData$Total.N~streamData$Distance)
abline(fit)
```


#function for resampling and fitting the data
#this function needs only one input, the dataframe that will be resampled and fitted.
#the function will resample the data, fit a lm, and then return the coefficient.
#This function is not actually run here, it is just defined.  It is run with replicate() below.
#It is best for the dataframe to have a different name inside the dataframe than it does elsewhere
#we define this name in the () in the function definition:
```{r}
resampleAndFit=function(thisStreamData){
  #within the function, streamData does not exist, but thisStreamData does
  
  #resample the response variable with replace = true
  thisStreamData$resampledResponse=sample(thisStreamData$Total.N,replace=T)
  
  #fit a lm to the resampled dataset, save it as resampleFit
  resampleFit=lm(thisStreamData$resampledResponse~thisStreamData$Distance)
  
  #points or fitted lines representing the resampled data can be added to the plot by uncommenting out these lines:
  #points(resampledResponse~thisStreamData$Distance,pch=2)
  #abline(resampleFit)
  
  #return the slope coefficient
 return(resampleFit$coefficients[[2]])
}

```

#run the command 'resampleAndFit(streamData)' 1000 times, and save the output as resampledSlopes
```{r}
resampledSlopes=replicate(10,resampleAndFit(streamData))
```

#is 1000 enough?  test by running a few times and seeing if the p value changes

#take a look at the distribution of the resampled slopes:

```{r}
hist(resampledSlopes)
```

#remember that in all the resampled data, there is no meaningful relationship between the x and y variables.
#therefore, the mean of the resampledSlopes should be close to 0.  If it's not, something is wrong

```{r}
mean(resampledSlopes)
```


#A p value as the probability that a pattern (e.g., slope) equivalent to 
# or more extreme than the observed pattern would occur if in fact there was no meaningful relationship.


# In this example, we are predicting a positive correlation (a positive slope).
# Therefore, we want to ask what proportion of the population of slopes fitted to the
# randomly resampled dataset are GREATER THAN the observed slope in the actual data
# the '>' operator asks weather resampledSlopes is greater than the actual fitted coefficient
# this returns a set of TRUE or FALSE values:
# TRUE where the resampled slope is greater than the slope of the actual data,
# and FALSE where the resampled slope is less than the slope of the actual data,
# R can also think of TRUE and FALSE as 1 and 0, thus the mean of the T/F set is the proportion of TRUEs

#therefore, this value is the p value:
```{r}
mean(resampledSlopes>fit$coefficients[2])
```
